Simulation Study: Varying Informativeness

### 2018-09-04 datasets/simulations/simulations-informative-cosine-nullAdaPT.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

In this set of simulations, we consider settings with both null and non-null
tests with an informative covariate. The covariate is sampled uniformly from
the interval [0, 1], and the conditional probability of a test being non-null
is a smooth (cosine) function of the covariate. We include simulation results
again with Gaussian, t-distributed, and Chi-Squared distributed noise.

@@

As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
for simulations with Gaussian or t-distributed noise. Again, we include both
`nulltype = "empirical"` and `nulltype = "theoretical"`.

@@

First, we consider the setting with Gaussian noise.

@@

Next, we consider the setting with t-distributed noise.

@@

For the t-distributed noise setting, we must specify the number of degrees of freedom
for ASH. We add an additional parameter to the `ashq` method with the corresponding
degrees of freedom of the noise distribution.

@@

Next, we consider the setting with t-distributed noise.

@@

Finally, we consider the setting with chi-squared distributed noise.

@@

For the chi-squared distributed noise setting, we must change the "mode" setting for ASH
from the default of 0 to `"estimate"` since the mode of null and alternative test statistics
are no longer centered at 0. While both approximate Normality and unimodality of effects are
violated in this simulation setting, by allowing the mode to be estimated, rather than forced
to 0, should return more comparable results.

>>>>

In this set of simulations, we consider settings with both null and non-null
tests with an informative covariate. The covariate is sampled uniformly from
the interval [0, 1], and the conditional probability of a test being non-null
is a smooth (cosine) function of the covariate. We include simulation results
again with Gaussian, t-distributed, and Chi-Squared distributed test statistics.

@@

As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
for simulations with Gaussian or t-distributed test statistics. Again, we include both
`nulltype = "empirical"` and `nulltype = "theoretical"`.

@@

First, we consider the setting with Gaussian test statistics.

@@

Next, we consider the setting with t-distributed test statistics.

@@

For the t-distributed setting, we must specify the number of degrees of freedom
for ASH. We add an additional parameter to the `ashq` method with the corresponding
degrees of freedom of the test statistic distribution.

@@

Next, we consider a second setting with t-distributed test statistics.

@@

Finally, we consider the setting with chi-squared distributed test statistics.

@@

For the chi-squared distributed setting, we must change the "mode" setting for ASH
from the default of 0 to `"estimate"` since the mode of null and alternative test statistics
are no longer centered at 0. While both approximate Normality and unimodality of effects are
violated in this simulation setting, by allowing the mode to be estimated, rather than forced
to 0, should return more comparable results.

### 2018-09-04 datasets/simulations/simulations-informative-cubic.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

In this set of simulations, we consider settings with both null and non-null
tests with an informative covariate. The covariate is sampled uniformly from
the interval [0, 1], and the conditional probability of a test being non-null
is a smooth (cubic) function of the covariate. We include simulation results
again with Gaussian, t-distributed, and Chi-Squared distributed noise.

@@

Next, we consider a second setting with t-distributed noise.

>>>>

In this set of simulations, we consider settings with both null and non-null
tests with an informative covariate. The covariate is sampled uniformly from
the interval [0, 1], and the conditional probability of a test being non-null
is a smooth (cubic) function of the covariate. We include simulation results
again with Gaussian, t-distributed, and Chi-Squared distributed test statistics.

@@

Next, we consider a second setting with t-distributed test statistics.

### 2018-09-04 datasets/simulations/simulations-null.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

First, we consider the null setting with gaussian noise.

@@

Next, we consider the null setting with t-distributed noise.

@@

Finally, we consider the null setting with chi-squared distributed noise.

>>>>

First, we consider the null setting with gaussian test statistics.

@@

Next, we consider the null setting with t-distributed test statistics.

@@

Finally, we consider the null setting with chi-squared distributed test statistics.

### 2018-09-04 datasets/simulations/simulations-varyingeffectsize-chisq.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

---
title: "Simulation Study: Varying Effect Size"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

In this set of simulations, we consider settings with both null and non-null
tests with varying distribution of effect sizes under the non-null (alternative)
setting. An informative covariate is included in the setting as described in
`simulations-informative-cubic.Rmd`. This set of simulations is similar to those
presented in `simulations-varyingeffectsize.Rmd`, however, Chi-Squared
distributed noise is used to perturb effects. Additionally, since the effect
must be non-negative (to be a valid value for the non-centrality parameter
of the a Chi-Squared distribution), the absolute value of all effect sizes
are used. Note that this modification still maintains a unimodal distribution
of the effect sizes for all settings except the "bimodal" case below.

>>>>

---
title: "Simulation Study: Varying Effect Size (chisq-4)"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

In this set of simulations, we consider settings with both null and non-null
tests with varying distribution of effect sizes under the non-null (alternative)
setting. An informative covariate is included in the setting as described in
`simulations-informative-cubic.Rmd`. This set of simulations is similar to those
presented in `simulations-varyingeffectsize.Rmd`, however, Chi-Squared
distributed test statistics are used. Additionally, since the effect
must be non-negative (to be a valid value for the non-centrality parameter
of the a Chi-Squared distribution), the absolute value of all effect sizes
are used. Note that this modification still maintains a unimodal distribution
of the effect sizes for all settings except the "bimodal" case below.

### 2018-09-04 datasets/simulations/simulations-varyingeffectsize-t.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

---
title: "Simulation Study: Varying Effect Size (t noise)"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

In this set of simulations, we consider settings with both null and non-null
tests with varying distribution of effect sizes under the non-null (alternative)
setting. An informative covariate is included in the setting as described in
`simulations-informative-cubic.Rmd`. This set of simulations is similar to those
presented in `simulations-varyingeffectsize.Rmd`, however, t-distributed noise
is used to perturb effects.

@@

As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
for simulations with Gaussian or t-distributed noise. Again, we include both
`nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
series of simulations use test statistics simulated with Gaussian noise, we include
Scott's FDR Regression in all of the comparisons.

@@

Since all simulation settings in this case study use t-distributed noise, we must specify
the number of degrees of freedom for ASH. We add an additional parameter to the `ashq`
method with the corresponding degrees of freedom of the noise distribution.

>>>>

---
title: "Simulation Study: Varying Effect Size (t-11)"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

@@

In this set of simulations, we consider settings with both null and non-null
tests with varying distribution of effect sizes under the non-null (alternative)
setting. An informative covariate is included in the setting as described in
`simulations-informative-cubic.Rmd`. This set of simulations is similar to those
presented in `simulations-varyingeffectsize.Rmd`, however, t-distributed test
statistics are used.

@@

As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
for simulations with Gaussian or t-distributed test statistics. Again, we include both
`nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
series of simulations use test statistics simulated with t-distributed test statistics,
we include Scott's FDR Regression in all of the comparisons.

@@

Since all simulation settings in this case study use t-distributed test statistics, we must specify
the number of degrees of freedom for ASH. We add an additional parameter to the `ashq`
method with the corresponding degrees of freedom of the test statistic distribution.

### 2018-09-04 datasets/simulations/simulations-varyinginformativeness-smooth.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
for simulations with Gaussian noise. Again, we include both
`nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
series of simulations use test statistics simulated with Gaussian noise, we include
Scott's FDR Regression in all of the comparisons.

>>>>

As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
for simulations with Gaussian test statistics. Again, we include both
`nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
series of simulations use test statistics simulated with Gaussian test statistics, we include
Scott's FDR Regression in all of the comparisons.

### 2018-09-04 datasets/simulations/simulations-varyinginformativeness.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

>>>>

First, we show the form of the relationship between the informative covariate and null
proportion, pi0, across varying levels of "informativeness."

```{r plot-informativeness, fig.width = 9, fig.height = 2.5}
xseq <- seq(0, 1, by = .01)
pi0trends <- lapply(seq(0, 1, by = .2), function(i) {
    tibble(info = i, x = xseq,
           pi0 = pi0_varyinfo80(i)(xseq))
})
pi0trends <- bind_rows(pi0trends)

ggplot(pi0trends, aes(x = x, y = pi0)) +
    geom_line() +
    geom_text(aes(label = round(pi0, 3), hjust = x, vjust = round(pi0)),
              dplyr::filter(pi0trends, x == 0 | x == 1),
              size = 3, color = 'blue') + 
    theme_bw() +
    scale_x_continuous("x (informative covariate)", breaks = 0:1) +
    scale_y_continuous(labels = scales::percent) +
    facet_grid(. ~ info, labeller = label_both) +
    ggtitle("Covariate vs. pi0 for varying informativeness")
```

Alternatively, since the function is a step function, we can simply plot the levels
of the two steps.

```{r, plot-informativeness-v2, fig.width = 6, fig.height = 5}
xseq <- seq(0, 1, by = .01)
pi0trends <- lapply(xseq, function(i) {
    tibble(info = i, x = 0:1,
           pi0 = pi0_varyinfo80(i)(0:1))
})
pi0trends <- bind_rows(pi0trends)
pi0trends <- dplyr::mutate(pi0trends,
                           x = factor(x, levels = 0:1,
                                      labels = c("[0, .8]", "(.8, 1]")))

ggplot(pi0trends, aes(x = info, y = pi0, group = x)) +
    geom_line() +
    geom_text(aes(label = x, vjust = round(pi0)),
              dplyr::filter(pi0trends, info == 1),
              hjust = 1, size = 3, color = 'blue') + 
    theme_bw() +
    scale_x_continuous("Informativeness", labels = scales::percent) +
    scale_y_continuous(labels = scales::percent) +
    ggtitle("null proportions for varying informativeness")
```
